### 2024-01-21 scripts/00-simulated_data.qmd
download the data, do the reference, and some introduction and data parts

---
title: "Simulated Data for the Analysis of Major Crimes in Toronto"
author: "Boxuan Yi"
format: html
---

```{r}
#### Preamble ####
# Purpose: Simulate the dataset of major crimes in Toronto
# Author: Boxuan Yi
# Email: boxuan.yi@mail.utoronto.ca
# Date: 17 January 2024
```