### 2024-01-23 scripts/00-simulated_data.qmd
Do the plan and simulation

---
title: "Simulated Data for the Analysis of Major Crimes in Toronto"
author: "Boxuan Yi"
format: pdf
---

```{r}
#| message: false
#| echo: false
#### Preamble ####
# Purpose: Simulate the dataset of major crimes in Toronto
# Author: Boxuan Yi
# Email: boxuan.yi@mail.utoronto.ca
# Date: 17 January 2024
```

```{r}
#| message: false
#| echo: false
# Plan
library(tidyverse)
library(janitor)

```

# Plan
![Plan Image](project/scripts/plan.png)

```{r}
# Simulate
#| message: false
#| echo: false
# Simulation for the number of crimes in each year (2014-2023) and the month with the most crimes in each year
set.seed(7)

months <- c("January", "February", "March", "April", "May", "June", "July",
            "August", "September", "October", "November", "December")

simulated_data1 <-
  tibble(
    "Year" = 2014:2023,
    "Number of Crimes" = runif(n = 10, min = 0, max = 10000),
    "Month" = sample(months, 10, replace = TRUE)
    )
simulated_data1

```

```{r}
#| message: false
#| echo: false
# Simulation for the number of crimes classified in major crime types
set.seed(7)

simulated_data2 <-
  tibble(
    "Year" = 2014:2023,
    "Number of Assault" = runif(n = 10, min = 0, max = 2000),
    "Number of Auto Theft" = runif(n = 10, min = 0, max = 2000),
    "Number of Break and Enter" = runif(n = 10, min = 0, max = 2000),
    "Number of Robbery" = runif(n = 10, min = 0, max = 2000),
    "Number of Theft Over" = runif(n = 10, min = 0, max = 2000)
    )
simulated_data2

```

```{r}
#| message: false
#| echo: false
# Simulation for the number of crimes classified in major crime types
set.seed(7)
simulated_data3 <-
  tibble(
    "Year" = 2014:2023,
    "Etobicoke" = runif(n = 10, min = 0, max = 1000),
    "Etobicoke-York" = runif(n = 10, min = 0, max = 1000),
    "North York" = runif(n = 10, min = 0, max = 1000),
    "Other" = runif(n = 10, min = 0, max = 1000),
    "Scarborough" = runif(n = 10, min = 0, max = 1000),
    "Toronto-East York" = runif(n = 10, min = 0, max = 1000)
    )
simulated_data3
```

```{r}
#| message: false
#| echo: false
# Simulation for the number of crimes classified in major crime types
set.seed(7)
simulated_data4 <-
  tibble(
    "Hour" = 0:23,
    "Etobicoke" = runif(n = 24, min = 0, max = 1000),
    )
simulated_data4
```