### 2024-01-23 script/01-download_data.qmd
Change the typo in scripts

---
title: "Download and Save the Actual Data on MCI"
author: "Boxuan Yi"
format: html
---

```{r}
#### Preamble ####
# Acquire
# Purpose: Download and save the actual data on major crimes that reported in Toronto 
# from 2014 to 2023
# Author: Boxuan Yi
# Email: boxuan.yi@mail.utoronto.ca
# Date: 18 January 2024
# Prerequisites: Know where to get major crimes data.
# I will use MCI as an abbreviation of Major Crime Indicator
```

```{r}
# Install all the packages I need for this analysis
install.packages('opendatatoronto')
install.packages('dplyr')
install.packages('tidyverse')
install.packages('janitor')
install.packages('leaflet')
install.packages('ggplot2')
install.packages('knitr')
install.packages('here')
```

```{r}
library(opendatatoronto)
library(dplyr)
library(tidyverse)
library(here)
library(janitor)
```

```{r}
# get package and all resources for this package
package <- show_package("major-crime-indicators")
resources <- list_package_resources("major-crime-indicators")

# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))

# load the second one
raw_mci_data <- filter(datastore_resources, row_number()==2) %>% get_resource()
```

```{r}
# Write data
write_csv (
  x = raw_mci_data,
  file = "unedited_data.csv"
)
```

```{r}
# Read and clean the dataset
raw_mci_data <- read_csv(
    file = here("inputs/data/unedited_data.csv"))
clean_mci_data <- clean_names(raw_mci_data)
```

```{r}
# Write the cleaned dataset
write_csv (
  x = clean_mci_data,
  file = "clean_mci.csv"
)
```

```{r}
clean_mci_data <- read_csv(
    file = here("inputs/data/clean_mci.csv"))
```